---
title: "Antithetic Variates"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🟡 Track 2: Intermediate</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 12 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Negative Covariance Principles
</div>

### 💡 Story-Driven Motivation
> Instead of drawing independent uniform numbers U1 and U2, pair U with its mirror reflection 1-U to balance out high and low outcomes!

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
library(srm)
# Run exported srm function:
antithetic_mc(g = function(x) exp(x), n = 5000)
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? Antithetic variates require zero extra random number generation—you get double the data for free!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Try modifying the parameters below:
antithetic_mc(g = function(x) exp(x), n = 5000)
```

```{r practice-ex-solution}
antithetic_mc(g = function(x) exp(x), n = 5000)
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think antithetic sampling works on non-monotonic functions. (Wrong! If g(x) is symmetric, Cov(g(U), g(1-U)) can be positive).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("Why does evaluating (g(U) + g(1-U))/2 reduce variance compared to independent uniform draws?",
    answer("Because Cov(g(U), g(1-U)) is negative for monotonic g", correct = ("Because Cov(g(U), g(1-U)) is negative for monotonic g" == "Because Cov(g(U), g(1-U)) is negative for monotonic g")),
    answer("Because it doubles the sample size", correct = ("Because it doubles the sample size" == "Because Cov(g(U), g(1-U)) is negative for monotonic g")),
    answer("Because 1-U is larger than U", correct = ("Because 1-U is larger than U" == "Because Cov(g(U), g(1-U)) is negative for monotonic g")),
    answer("Because it eliminates bias", correct = ("Because it eliminates bias" == "Because Cov(g(U), g(1-U)) is negative for monotonic g")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
