---
title: "Antithetic Variates"
description: "Reduce Monte Carlo integration variance using negative correlation pairing (U, 1-U) from scratch."
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🟡 Track 2: Intermediate</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 12 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Negative Covariance Principles
</div>

### 💡 Story-Driven Motivation
> Instead of drawing independent uniform numbers U1 and U2, pair U with its mirror reflection 1-U to balance out high and low outcomes!

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
# Step-by-Step Antithetic Variates from Scratch
n_pairs <- 2500 # 2500 pairs = 5000 evaluations
g <- function(x) exp(x)

u <- runif(n_pairs)
# Pair U with mirror 1 - U
pair_evals <- (g(u) + g(1 - u)) / 2

u_raw <- runif(5000)
raw_evals <- g(u_raw)

cat("Antithetic Estimate:", mean(pair_evals), " (Var:", var(pair_evals), ")
")
cat("Raw MC Estimate:", mean(raw_evals), " (Var:", var(raw_evals), ")
")
cat("Variance Reduction:", round(100 * (1 - var(pair_evals)/var(raw_evals)), 2), "%
")
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? Antithetic variates require zero extra random number generation—you get double the data for free!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>g</code></td><td><code>function(x) exp(x)</code></td><td>Monotonic integrand function g(x) on interval (0,1).</td></tr>
  <tr><td><code>n</code></td><td><code>5000</code></td><td>Total sample size (split into n/2 antithetic pairs (U, 1-U)).</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?antithetic_mc</code> or <code>help(antithetic_mc, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Step-by-Step Antithetic Variates from Scratch
n_pairs <- 2500 # 2500 pairs = 5000 evaluations
g <- function(x) exp(x)

u <- runif(n_pairs)
# Pair U with mirror 1 - U
pair_evals <- (g(u) + g(1 - u)) / 2

u_raw <- runif(5000)
raw_evals <- g(u_raw)

cat("Antithetic Estimate:", mean(pair_evals), " (Var:", var(pair_evals), ")
")
cat("Raw MC Estimate:", mean(raw_evals), " (Var:", var(raw_evals), ")
")
cat("Variance Reduction:", round(100 * (1 - var(pair_evals)/var(raw_evals)), 2), "%
")
```

```{r practice-ex-solution}
# Step-by-Step Antithetic Variates from Scratch
n_pairs <- 2500 # 2500 pairs = 5000 evaluations
g <- function(x) exp(x)

u <- runif(n_pairs)
# Pair U with mirror 1 - U
pair_evals <- (g(u) + g(1 - u)) / 2

u_raw <- runif(5000)
raw_evals <- g(u_raw)

cat("Antithetic Estimate:", mean(pair_evals), " (Var:", var(pair_evals), ")
")
cat("Raw MC Estimate:", mean(raw_evals), " (Var:", var(raw_evals), ")
")
cat("Variance Reduction:", round(100 * (1 - var(pair_evals)/var(raw_evals)), 2), "%
")
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think antithetic sampling works on non-monotonic functions. (Wrong! If g(x) is symmetric, Cov(g(U), g(1-U)) can be positive).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("Why does evaluating (g(U) + g(1-U))/2 reduce variance compared to independent uniform draws?",
    answer("Because Cov(g(U), g(1-U)) is negative for monotonic g", correct = ("Because Cov(g(U), g(1-U)) is negative for monotonic g" == "Because Cov(g(U), g(1-U)) is negative for monotonic g")),
    answer("Because it doubles the sample size", correct = ("Because it doubles the sample size" == "Because Cov(g(U), g(1-U)) is negative for monotonic g")),
    answer("Because 1-U is larger than U", correct = ("Because 1-U is larger than U" == "Because Cov(g(U), g(1-U)) is negative for monotonic g")),
    answer("Because it eliminates bias", correct = ("Because it eliminates bias" == "Because Cov(g(U), g(1-U)) is negative for monotonic g")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
