---
title: "Multi-Level Splitting"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🔴 Track 3: Advanced</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 18 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Rare Event Level Thresholds
</div>

### 💡 Story-Driven Motivation
> Estimating rare cascade failures in power grids or financial networks by splitting successful particles across intermediate level thresholds L1 < L2 < ... < Lm.

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
library(srm)
# Run exported srm function:
multilevel_splitting(levels = c(2, 4, 6), n_particles = 1000)
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? Multi-level splitting multiplies surviving particles at each threshold, turning a 1-in-a-billion event into feasible sub-problems!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Try modifying the parameters below:
multilevel_splitting(levels = c(2, 4, 6), n_particles = 1000)
```

```{r practice-ex-solution}
multilevel_splitting(levels = c(2, 4, 6), n_particles = 1000)
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually choose thresholds too far apart. (Ideal thresholds keep intermediate step probabilities around 0.1 to 0.3).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("How does Multi-Level Splitting calculate the final rare event probability P(A)?",
    answer("By multiplying conditional probabilities across intermediate levels P(A_k | A_{k-1})", correct = ("By multiplying conditional probabilities across intermediate levels P(A_k | A_{k-1})" == "By multiplying conditional probabilities across intermediate levels P(A_k | A_{k-1})")),
    answer("By adding all level counts", correct = ("By adding all level counts" == "By multiplying conditional probabilities across intermediate levels P(A_k | A_{k-1})")),
    answer("By taking the maximum probability", correct = ("By taking the maximum probability" == "By multiplying conditional probabilities across intermediate levels P(A_k | A_{k-1})")),
    answer("By average", correct = ("By average" == "By multiplying conditional probabilities across intermediate levels P(A_k | A_{k-1})")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
