---
title: "Metropolis-Hastings Algorithm"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🔴 Track 3: Advanced</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 20 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Markov Chains & Proposal Distributions
</div>

### 💡 Story-Driven Motivation
> How do statisticians sample from complex 100-dimensional Bayesian posterior distributions when formulas are impossible to integrate? The Metropolis-Hastings algorithm steps through the landscape with random-walk candidate moves!

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
library(srm)
# Run exported srm function:
metropolis_hastings(target_log_density = function(x) dnorm(x, log=TRUE), init = 0, n_iter = 5000, proposal_sd = 1)
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? Metropolis-Hastings was chosen by IEEE Computing as one of the top 10 most influential algorithms of the 20th century!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Try modifying the parameters below:
metropolis_hastings(target_log_density = function(x) dnorm(x, log=TRUE), init = 0, n_iter = 5000, proposal_sd = 1)
```

```{r practice-ex-solution}
metropolis_hastings(target_log_density = function(x) dnorm(x, log=TRUE), init = 0, n_iter = 5000, proposal_sd = 1)
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think MCMC samples are independent random numbers. (Wrong! MCMC samples are autocorrelated Markov steps).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("In high-dimensional target spaces, what is the theoretical optimal MCMC target acceptance rate for random-walk Metropolis?",
    answer("Approximately 23.4%", correct = ("Approximately 23.4%" == "Approximately 23.4%")),
    answer("100%", correct = ("100%" == "Approximately 23.4%")),
    answer("50%", correct = ("50%" == "Approximately 23.4%")),
    answer("5.0%", correct = ("5.0%" == "Approximately 23.4%")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
