---
title: "Gibbs Sampling"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🔴 Track 3: Advanced</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 20 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Full Conditional Distributions
</div>

### 💡 Story-Driven Motivation
> Instead of proposing multi-dimensional jumps, Gibbs sampling updates one variable at a time using its full conditional distribution pi(xi | x_-i) with a 100% acceptance rate!

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
library(srm)
# Run exported srm function:
gibbs_sampler(n_iter = 5000, rho = 0.8)
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? Gibbs sampling was named after physicist Josiah Willard Gibbs by brothers Stuart and Donald Geman in 1984!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>n_iter</code></td><td><code>5000</code></td><td>Total Gibbs sampling iterations.</td></tr>
  <tr><td><code>rho</code></td><td><code>0.8</code></td><td>Correlation parameter rho in (-1, 1) between X and Y.</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?gibbs_sampler</code> or <code>help(gibbs_sampler, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Try modifying the parameters below:
gibbs_sampler(n_iter = 5000, rho = 0.8)
```

```{r practice-ex-solution}
gibbs_sampler(n_iter = 5000, rho = 0.8)
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think Gibbs sampler steps can be rejected. (Wrong! Gibbs sampler acceptance ratio alpha is always identically 1).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("What is the acceptance probability alpha for candidate moves in a standard Gibbs sampler?",
    answer("Always 1.0 (100% acceptance)", correct = ("Always 1.0 (100% acceptance)" == "Always 1.0 (100% acceptance)")),
    answer("Depends on proposal SD", correct = ("Depends on proposal SD" == "Always 1.0 (100% acceptance)")),
    answer("0.234", correct = ("0.234" == "Always 1.0 (100% acceptance)")),
    answer("0.50", correct = ("0.50" == "Always 1.0 (100% acceptance)")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
