---
title: "Robbins-Monro Stochastic Approximation"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🔴 Track 3: Advanced</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 15 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Stochastic Root-Finding & Step Decay
</div>

### 💡 Story-Driven Motivation
> Finding the root of a noisy, fluctuating function without computing explicit derivatives. Herbert Robbins and Sutton Monro pioneered stochastic approximation in 1951.

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
library(srm)
# Run exported srm function:
robbins_monro(noisy_g = function(th) (th - 3) + rnorm(1, sd=0.2), init = 0, n_iter = 1000, c = 1)
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? Robbins-Monro is the direct ancestor of Stochastic Gradient Descent (SGD) used to train deep neural networks!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>noisy_g</code></td><td><code>function(th) (th - 3) + rnorm(1, sd=0.2)</code></td><td>Function evaluating noisy measurement g(theta) + noise.</td></tr>
  <tr><td><code>init</code></td><td><code>0</code></td><td>Initial starting guess theta_0.</td></tr>
  <tr><td><code>n_iter</code></td><td><code>1000</code></td><td>Iterations.</td></tr>
  <tr><td><code>c</code></td><td><code>1</code></td><td>Step size scale constant for step decay a_n = c/n.</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?robbins_monro</code> or <code>help(robbins_monro, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Try modifying the parameters below:
robbins_monro(noisy_g = function(th) (th - 3) + rnorm(1, sd=0.2), init = 0, n_iter = 1000, c = 1)
```

```{r practice-ex-solution}
robbins_monro(noisy_g = function(th) (th - 3) + rnorm(1, sd=0.2), init = 0, n_iter = 1000, c = 1)
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually keep step size a_n constant. (Wrong! Step sizes must decay such that sum(a_n) = infinity and sum(a_n^2) < infinity).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("Which step size decay sequence a_n satisfies the Robbins-Monro convergence conditions sum(a_n) = infinity and sum(a_n^2) < infinity?",
    answer("a_n = 1 / n", correct = ("a_n = 1 / n" == "a_n = 1 / n")),
    answer("a_n = 1 / n^2", correct = ("a_n = 1 / n^2" == "a_n = 1 / n")),
    answer("a_n = 0.1 (constant)", correct = ("a_n = 0.1 (constant)" == "a_n = 1 / n")),
    answer("a_n = 1 / sqrt(n)", correct = ("a_n = 1 / sqrt(n)" == "a_n = 1 / n")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
