---
title: "Inverse-Transform Method"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🟢 Track 1: Essentials</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 15 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Calculus Cumulative Distribution Functions (CDF)
</div>

### 💡 Story-Driven Motivation
> If a computer can only generate uniform numbers U ~ Unif(0,1), how do we sample exponential wait times or heavy-tailed financial returns? We pass U through the inverse CDF: X = F^-1(U).

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
library(srm)
# Run exported srm function:
inv_transform_exp(n = 100, rate = 2)
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? Inverse-transform is exact—it maps 100% of probability mass without rejecting a single draw!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>n</code></td><td><code>100</code></td><td>Number of exponential random variates to generate.</td></tr>
  <tr><td><code>rate</code></td><td><code>2</code></td><td>Rate parameter lambda > 0 of the Exponential distribution (mean = 1/lambda).</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?inv_transform_exp</code> or <code>help(inv_transform_exp, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Try modifying the parameters below:
inv_transform_exp(n = 100, rate = 2)
```

```{r practice-ex-solution}
inv_transform_exp(n = 100, rate = 2)
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think U and 1-U give different distributions. (Wrong! Since U is symmetric on (0,1), 1-U is also Unif(0,1)).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("If X ~ Exp(lambda) with CDF F(x) = 1 - exp(-lambda * x), what is the inverse CDF F^-1(u)?",
    answer("-log(1-u) / lambda", correct = ("-log(1-u) / lambda" == "-log(1-u) / lambda")),
    answer("exp(u) * lambda", correct = ("exp(u) * lambda" == "-log(1-u) / lambda")),
    answer("sqrt(u)", correct = ("sqrt(u)" == "-log(1-u) / lambda")),
    answer("1 / (1-u)", correct = ("1 / (1-u)" == "-log(1-u) / lambda")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
