---
title: "Sampling Designs & Power Calculations"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🟢 Track 1: Essentials</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 12 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Basic Probability & Confidence Intervals
</div>

### 💡 Story-Driven Motivation
> In a regional healthcare study, public health officials needed to survey 1,200 households to infer health coverage for 5 million citizens without bias. How did they calculate sample size and allocate samples across strata?

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
library(srm)
# Run exported srm function:
calc_sample_size(p = 0.5, d = 0.05, conf_level = 0.95)
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? The 1936 Literary Digest poll surveyed 2.4 million people but wrongly predicted the US election because of sampling bias!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Try modifying the parameters below:
calc_sample_size(p = 0.5, d = 0.05, conf_level = 0.95)
```

```{r practice-ex-solution}
calc_sample_size(p = 0.5, d = 0.05, conf_level = 0.95)
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think bigger samples compensate for selection bias. (Wrong! A biased sample of 1 million is still a biased sample).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("What happens to the required sample size n if you cut the margin of error d in half?",
    answer("It doubles (2x)", correct = ("It doubles (2x)" == "It quadruples (4x)")),
    answer("It quadruples (4x)", correct = ("It quadruples (4x)" == "It quadruples (4x)")),
    answer("It stays the same", correct = ("It stays the same" == "It quadruples (4x)")),
    answer("It decreases by half", correct = ("It decreases by half" == "It quadruples (4x)")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
