---
title: "Sampling Designs & Power Calculations"
description: "Learn sample size formulas, Neyman stratified allocation, and CONSORT trial design flowcharts in R."
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🟢 Track 1: Essentials</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 12 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Basic Probability & Confidence Intervals
</div>

### 💡 Story-Driven Motivation
> In a regional healthcare study, public health officials needed to survey 1,200 households to infer health coverage for 5 million citizens without bias. How did they calculate sample size and allocate samples across strata?

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
# Step-by-Step Sample Size Calculation for Proportion from Scratch
p <- 0.5       # Anticipated population proportion
d <- 0.05      # Margin of error (precision bound)
conf <- 0.95   # Confidence level

# Calculate Z-score for 95% confidence (Z = 1.96)
alpha <- 1 - conf
z_score <- qnorm(1 - alpha / 2)

# Apply formula: n = (Z^2 * p * (1 - p)) / d^2
n_raw <- (z_score^2 * p * (1 - p)) / (d^2)
n_final <- ceiling(n_raw) # Round up to whole integer

cat("Z-score:", z_score, "
")
cat("Exact required n:", n_raw, "
")
cat("Final sample size (rounded up):", n_final, "
")
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? The 1936 Literary Digest poll surveyed 2.4 million people but wrongly predicted the US election because of sampling bias!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>p</code></td><td><code>0.5</code></td><td>Anticipated population proportion or prevalence. Setting p = 0.5 is conservative as it yields maximum variance p(1-p).</td></tr>
  <tr><td><code>d</code></td><td><code>0.05</code></td><td>Desired margin of error or precision bound (e.g., 0.05 means ±5% precision).</td></tr>
  <tr><td><code>conf_level</code></td><td><code>0.95</code></td><td>Confidence level (e.g., 0.95 corresponds to a 95% confidence interval where Z = 1.96).</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?calc_sample_size</code> or <code>help(calc_sample_size, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Step-by-Step Sample Size Calculation for Proportion from Scratch
p <- 0.5       # Anticipated population proportion
d <- 0.05      # Margin of error (precision bound)
conf <- 0.95   # Confidence level

# Calculate Z-score for 95% confidence (Z = 1.96)
alpha <- 1 - conf
z_score <- qnorm(1 - alpha / 2)

# Apply formula: n = (Z^2 * p * (1 - p)) / d^2
n_raw <- (z_score^2 * p * (1 - p)) / (d^2)
n_final <- ceiling(n_raw) # Round up to whole integer

cat("Z-score:", z_score, "
")
cat("Exact required n:", n_raw, "
")
cat("Final sample size (rounded up):", n_final, "
")
```

```{r practice-ex-solution}
# Step-by-Step Sample Size Calculation for Proportion from Scratch
p <- 0.5       # Anticipated population proportion
d <- 0.05      # Margin of error (precision bound)
conf <- 0.95   # Confidence level

# Calculate Z-score for 95% confidence (Z = 1.96)
alpha <- 1 - conf
z_score <- qnorm(1 - alpha / 2)

# Apply formula: n = (Z^2 * p * (1 - p)) / d^2
n_raw <- (z_score^2 * p * (1 - p)) / (d^2)
n_final <- ceiling(n_raw) # Round up to whole integer

cat("Z-score:", z_score, "
")
cat("Exact required n:", n_raw, "
")
cat("Final sample size (rounded up):", n_final, "
")
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think bigger samples compensate for selection bias. (Wrong! A biased sample of 1 million is still a biased sample).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("What happens to the required sample size n if you cut the margin of error d in half?",
    answer("It doubles (2x)", correct = ("It doubles (2x)" == "It quadruples (4x)")),
    answer("It quadruples (4x)", correct = ("It quadruples (4x)" == "It quadruples (4x)")),
    answer("It stays the same", correct = ("It stays the same" == "It quadruples (4x)")),
    answer("It decreases by half", correct = ("It decreases by half" == "It quadruples (4x)")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
