---
title: "Poisson Processes & Arrival Times"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🟡 Track 2: Intermediate</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 12 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Exponential Inter-Arrival Distributions
</div>

### 💡 Story-Driven Motivation
> How do emergency rooms model patient surge patterns over a 24-hour shift? By simulating Poisson point processes with exponential inter-arrival times.

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
library(srm)
# Run exported srm function:
sim_poisson_process(T_max = 10, rate = 2)
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? Simon Poisson originally derived his distribution while studying horse-kick fatalities in the Prussian army!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>T_max</code></td><td><code>10</code></td><td>Total simulation time horizon.</td></tr>
  <tr><td><code>rate</code></td><td><code>2</code></td><td>Event arrival rate lambda > 0 (average events per unit time).</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?sim_poisson_process</code> or <code>help(sim_poisson_process, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Try modifying the parameters below:
sim_poisson_process(T_max = 10, rate = 2)
```

```{r practice-ex-solution}
sim_poisson_process(T_max = 10, rate = 2)
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think Poisson arrival times are fixed. (Wrong! Inter-arrival times are memoryless Exponential random variables).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("If event arrivals follow a Poisson process with rate lambda, what is the distribution of the time between consecutive events?",
    answer("Exponential(rate = lambda)", correct = ("Exponential(rate = lambda)" == "Exponential(rate = lambda)")),
    answer("Uniform(0, lambda)", correct = ("Uniform(0, lambda)" == "Exponential(rate = lambda)")),
    answer("Normal(lambda, 1)", correct = ("Normal(lambda, 1)" == "Exponential(rate = lambda)")),
    answer("Binomial(lambda)", correct = ("Binomial(lambda)" == "Exponential(rate = lambda)")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
