---
title: "Non-Parametric Bootstrap Resampling"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🟢 Track 1: Essentials</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 15 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Sampling Distributions & Standard Error
</div>

### 💡 Story-Driven Motivation
> How do you calculate a 95% confidence interval for median patient hospital stay lengths when you don't know the underlying distribution? Pull yourself up by your own bootstraps!

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
# Step-by-Step Non-Parametric Bootstrap Resampling from Scratch
data_sample <- c(23, 45, 67, 34, 56, 78, 89, 90, 45, 67)
R_replicates <- 1000
n_obs <- length(data_sample)

boot_means <- numeric(R_replicates)

# Resampling Loop with Replacement
for (i in 1:R_replicates) {
  boot_draw <- sample(data_sample, size = n_obs, replace = TRUE)
  boot_means[i] <- mean(boot_draw)
}

# Standard error and 95% Percentile Confidence Interval
boot_se <- sd(boot_means)
ci_95 <- quantile(boot_means, probs = c(0.025, 0.975))

cat("Original Sample Mean:", mean(data_sample), "
")
cat("Bootstrap SE:", boot_se, "
")
cat("95% Percentile CI:", ci_95, "
")
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? Bradley Efron introduced the Bootstrap in 1979, revolutionizing applied statistics without assuming bell curves!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>data</code></td><td><code>ehr_cohort$age</code></td><td>Numeric vector of observed sample data.</td></tr>
  <tr><td><code>statistic</code></td><td><code>mean</code></td><td>Function computing the parameter statistic of interest (e.g., mean, median).</td></tr>
  <tr><td><code>R</code></td><td><code>1000</code></td><td>Number of bootstrap resamples drawn with replacement.</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?bootstrap_ci</code> or <code>help(bootstrap_ci, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Step-by-Step Non-Parametric Bootstrap Resampling from Scratch
data_sample <- c(23, 45, 67, 34, 56, 78, 89, 90, 45, 67)
R_replicates <- 1000
n_obs <- length(data_sample)

boot_means <- numeric(R_replicates)

# Resampling Loop with Replacement
for (i in 1:R_replicates) {
  boot_draw <- sample(data_sample, size = n_obs, replace = TRUE)
  boot_means[i] <- mean(boot_draw)
}

# Standard error and 95% Percentile Confidence Interval
boot_se <- sd(boot_means)
ci_95 <- quantile(boot_means, probs = c(0.025, 0.975))

cat("Original Sample Mean:", mean(data_sample), "
")
cat("Bootstrap SE:", boot_se, "
")
cat("95% Percentile CI:", ci_95, "
")
```

```{r practice-ex-solution}
# Step-by-Step Non-Parametric Bootstrap Resampling from Scratch
data_sample <- c(23, 45, 67, 34, 56, 78, 89, 90, 45, 67)
R_replicates <- 1000
n_obs <- length(data_sample)

boot_means <- numeric(R_replicates)

# Resampling Loop with Replacement
for (i in 1:R_replicates) {
  boot_draw <- sample(data_sample, size = n_obs, replace = TRUE)
  boot_means[i] <- mean(boot_draw)
}

# Standard error and 95% Percentile Confidence Interval
boot_se <- sd(boot_means)
ci_95 <- quantile(boot_means, probs = c(0.025, 0.975))

cat("Original Sample Mean:", mean(data_sample), "
")
cat("Bootstrap SE:", boot_se, "
")
cat("95% Percentile CI:", ci_95, "
")
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think Bootstrap creates brand-new data points. (Wrong! It resamples your existing dataset with replacement).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("In non-parametric bootstrap resampling of sample size n, how are bootstrap samples drawn from the original data?",
    answer("With replacement, size n", correct = ("With replacement, size n" == "With replacement, size n")),
    answer("Without replacement, size n/2", correct = ("Without replacement, size n/2" == "With replacement, size n")),
    answer("Sorted by rank", correct = ("Sorted by rank" == "With replacement, size n")),
    answer("From a Normal distribution", correct = ("From a Normal distribution" == "With replacement, size n")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
