---
title: "Coupling From The Past (CFTP)"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🔴 Track 3: Advanced</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 15 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Monotonic Markov Chains
</div>

### 💡 Story-Driven Motivation
> Standard MCMC requires burn-in estimation. James Propp and David Wilson invented CFTP in 1996 to produce 100% exact, un-biased samples by running chains backward in time!

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
library(srm)
# Run exported srm function:
cftp_perfect_sampler(max_lookback = 100)
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? CFTP is called 'perfect sampling' because it completely eliminates burn-in bias!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>max_lookback</code></td><td><code>100</code></td><td>Maximum backward time horizon -T to check for chain coalescence.</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?cftp_perfect_sampler</code> or <code>help(cftp_perfect_sampler, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Try modifying the parameters below:
cftp_perfect_sampler(max_lookback = 100)
```

```{r practice-ex-solution}
cftp_perfect_sampler(max_lookback = 100)
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually try running CFTP forward in time. (Wrong! Forward coupling leads to readout bias. CFTP MUST start in the past).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("Why must Coupling From The Past (CFTP) initialize chains at time -T in the past rather than forward from time 0?",
    answer("To prevent state selection readout bias at time 0", correct = ("To prevent state selection readout bias at time 0" == "To prevent state selection readout bias at time 0")),
    answer("To make computation faster", correct = ("To make computation faster" == "To prevent state selection readout bias at time 0")),
    answer("Because time flows backward in MCMC", correct = ("Because time flows backward in MCMC" == "To prevent state selection readout bias at time 0")),
    answer("To avoid random numbers", correct = ("To avoid random numbers" == "To prevent state selection readout bias at time 0")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
