---
title: "Coupling From The Past (CFTP)"
description: "Implement Coupling From The Past (CFTP) exact un-biased MCMC sampling without burn-in."
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🔴 Track 3: Advanced</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 15 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Monotonic Markov Chains
</div>

### 💡 Story-Driven Motivation
> Standard MCMC requires burn-in estimation. James Propp and David Wilson invented CFTP in 1996 to produce 100% exact, un-biased samples by running chains backward in time!

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
# Step-by-Step Coupling From The Past (CFTP) Monotonic Sampler from Scratch
max_lookback <- 64
t_back <- 1
coalesced <- FALSE
final_state <- NA

while (!coalesced && t_back <= max_lookback) {
  # Fixed random draws for this backward horizon
  u_randoms <- runif(t_back)
  
  state_low <- 0
  state_high <- 1
  
  for (step in 1:t_back) {
    u <- u_randoms[step]
    state_low <- ifelse(u < 0.5, max(0, state_low - 1), min(1, state_low + 1))
    state_high <- ifelse(u < 0.5, max(0, state_high - 1), min(1, state_high + 1))
  }
  
  if (state_low == state_high) {
    coalesced <- TRUE
    final_state <- state_low
  } else {
    t_back <- t_back * 2
  }
}

cat("Coalesced:", coalesced, "at lookback time T =", t_back, "
")
cat("Exact Exact Sample Value:", final_state, "
")
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? CFTP is called 'perfect sampling' because it completely eliminates burn-in bias!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>max_lookback</code></td><td><code>100</code></td><td>Maximum backward time horizon -T to check for chain coalescence.</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?cftp_perfect_sampler</code> or <code>help(cftp_perfect_sampler, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Step-by-Step Coupling From The Past (CFTP) Monotonic Sampler from Scratch
max_lookback <- 64
t_back <- 1
coalesced <- FALSE
final_state <- NA

while (!coalesced && t_back <= max_lookback) {
  # Fixed random draws for this backward horizon
  u_randoms <- runif(t_back)
  
  state_low <- 0
  state_high <- 1
  
  for (step in 1:t_back) {
    u <- u_randoms[step]
    state_low <- ifelse(u < 0.5, max(0, state_low - 1), min(1, state_low + 1))
    state_high <- ifelse(u < 0.5, max(0, state_high - 1), min(1, state_high + 1))
  }
  
  if (state_low == state_high) {
    coalesced <- TRUE
    final_state <- state_low
  } else {
    t_back <- t_back * 2
  }
}

cat("Coalesced:", coalesced, "at lookback time T =", t_back, "
")
cat("Exact Exact Sample Value:", final_state, "
")
```

```{r practice-ex-solution}
# Step-by-Step Coupling From The Past (CFTP) Monotonic Sampler from Scratch
max_lookback <- 64
t_back <- 1
coalesced <- FALSE
final_state <- NA

while (!coalesced && t_back <= max_lookback) {
  # Fixed random draws for this backward horizon
  u_randoms <- runif(t_back)
  
  state_low <- 0
  state_high <- 1
  
  for (step in 1:t_back) {
    u <- u_randoms[step]
    state_low <- ifelse(u < 0.5, max(0, state_low - 1), min(1, state_low + 1))
    state_high <- ifelse(u < 0.5, max(0, state_high - 1), min(1, state_high + 1))
  }
  
  if (state_low == state_high) {
    coalesced <- TRUE
    final_state <- state_low
  } else {
    t_back <- t_back * 2
  }
}

cat("Coalesced:", coalesced, "at lookback time T =", t_back, "
")
cat("Exact Exact Sample Value:", final_state, "
")
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually try running CFTP forward in time. (Wrong! Forward coupling leads to readout bias. CFTP MUST start in the past).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("Why must Coupling From The Past (CFTP) initialize chains at time -T in the past rather than forward from time 0?",
    answer("To prevent state selection readout bias at time 0", correct = ("To prevent state selection readout bias at time 0" == "To prevent state selection readout bias at time 0")),
    answer("To make computation faster", correct = ("To make computation faster" == "To prevent state selection readout bias at time 0")),
    answer("Because time flows backward in MCMC", correct = ("Because time flows backward in MCMC" == "To prevent state selection readout bias at time 0")),
    answer("To avoid random numbers", correct = ("To avoid random numbers" == "To prevent state selection readout bias at time 0")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
