---
title: "Accept-Reject Sampling"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🟡 Track 2: Intermediate</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 15 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Probability Density Functions & Envelopes
</div>

### 💡 Story-Driven Motivation
> What do you do when an inverse CDF F^-1(u) is impossible to solve analytically? Draw candidates from a proposal density g(x) under an envelope c*g(x), and accept or reject!

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
# Step-by-Step Accept-Reject Sampler for Beta(2, 5) from Scratch
n_target <- 500
shape1 <- 2
shape2 <- 5

# Compute bounding constant c = max f(x)
mode_x <- (shape1 - 1) / (shape1 + shape2 - 2)
c_envelope <- dbeta(mode_x, shape1, shape2)

samples <- numeric(n_target)
count <- 0
attempts <- 0

while (count < n_target) {
  attempts <- attempts + 1
  y <- runif(1) # Proposal candidate from Unif(0,1)
  u <- runif(1) # Decision uniform draw
  
  # Acceptance decision condition: U <= f(y) / (c * g(y))
  if (u <= dbeta(y, shape1, shape2) / c_envelope) {
    count <- count + 1
    samples[count] <- y
  }
}

cat("Acceptance Efficiency:", count / attempts, "
")
hist(samples, probability = TRUE, main = "Accept-Reject Beta(2,5) Samples")
curve(dbeta(x, shape1, shape2), add = TRUE, col = "red", lwd = 2)
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? Accept-reject is the secret engine inside R's rbeta() and rgamma() generators!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>n</code></td><td><code>500</code></td><td>Target number of accepted samples desired.</td></tr>
  <tr><td><code>shape1</code></td><td><code>2</code></td><td>Alpha shape parameter of target Beta(alpha, beta) distribution.</td></tr>
  <tr><td><code>shape2</code></td><td><code>5</code></td><td>Beta shape parameter of target Beta(alpha, beta) distribution.</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?accept_reject_beta</code> or <code>help(accept_reject_beta, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Step-by-Step Accept-Reject Sampler for Beta(2, 5) from Scratch
n_target <- 500
shape1 <- 2
shape2 <- 5

# Compute bounding constant c = max f(x)
mode_x <- (shape1 - 1) / (shape1 + shape2 - 2)
c_envelope <- dbeta(mode_x, shape1, shape2)

samples <- numeric(n_target)
count <- 0
attempts <- 0

while (count < n_target) {
  attempts <- attempts + 1
  y <- runif(1) # Proposal candidate from Unif(0,1)
  u <- runif(1) # Decision uniform draw
  
  # Acceptance decision condition: U <= f(y) / (c * g(y))
  if (u <= dbeta(y, shape1, shape2) / c_envelope) {
    count <- count + 1
    samples[count] <- y
  }
}

cat("Acceptance Efficiency:", count / attempts, "
")
hist(samples, probability = TRUE, main = "Accept-Reject Beta(2,5) Samples")
curve(dbeta(x, shape1, shape2), add = TRUE, col = "red", lwd = 2)
```

```{r practice-ex-solution}
# Step-by-Step Accept-Reject Sampler for Beta(2, 5) from Scratch
n_target <- 500
shape1 <- 2
shape2 <- 5

# Compute bounding constant c = max f(x)
mode_x <- (shape1 - 1) / (shape1 + shape2 - 2)
c_envelope <- dbeta(mode_x, shape1, shape2)

samples <- numeric(n_target)
count <- 0
attempts <- 0

while (count < n_target) {
  attempts <- attempts + 1
  y <- runif(1) # Proposal candidate from Unif(0,1)
  u <- runif(1) # Decision uniform draw
  
  # Acceptance decision condition: U <= f(y) / (c * g(y))
  if (u <= dbeta(y, shape1, shape2) / c_envelope) {
    count <- count + 1
    samples[count] <- y
  }
}

cat("Acceptance Efficiency:", count / attempts, "
")
hist(samples, probability = TRUE, main = "Accept-Reject Beta(2,5) Samples")
curve(dbeta(x, shape1, shape2), add = TRUE, col = "red", lwd = 2)
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think a higher envelope constant c is better. (Wrong! Higher c means lower efficiency 1/c and more rejected samples).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("What is the expected acceptance probability (efficiency) of the Accept-Reject algorithm with bounding constant c?",
    answer("1 / c", correct = ("1 / c" == "1 / c")),
    answer("c / 2", correct = ("c / 2" == "1 / c")),
    answer("1 - c", correct = ("1 - c" == "1 / c")),
    answer("exp(-c)", correct = ("exp(-c)" == "1 / c")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
