---
title: "Importance Sampling for Rare Events"
description: "Estimate rare event probabilities and extreme tail risks using importance sampling proposal densities."
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🟡 Track 2: Intermediate</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 18 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Proposal Densities & Weight Ratios
</div>

### 💡 Story-Driven Motivation
> NASA needs to estimate a 1-in-a-million structural failure probability. Ordinary Monte Carlo would require 100 million runs. Importance sampling shifts the proposal density g(x) to sample directly in the failure zone!

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
# Step-by-Step Importance Sampling for Rare Event (X > 3) from Scratch
n <- 5000
# Target density f(x) = N(0,1)
# Proposal density g(x) = N(3,1) centered at rare event tail
target_f <- function(x) dnorm(x, mean = 0, sd = 1)
proposal_g <- function(x) dnorm(x, mean = 3, sd = 1)

# Draw candidates from proposal g(x)
x_proposal <- rnorm(n, mean = 3, sd = 1)

# Compute importance weights w(x) = f(x) / g(x)
w <- target_f(x_proposal) / proposal_g(x_proposal)
evals <- (x_proposal > 3) * w

is_estimate <- mean(evals)
is_se <- sd(evals) / sqrt(n)

cat("True P(X > 3):", 1 - pnorm(3), "
")
cat("Importance Sampling Estimate:", is_estimate, "
")
cat("Standard Error:", is_se, "
")
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? Importance sampling is widely used in nuclear shielding simulations to track rare particle penetration!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>h</code></td><td><code>function(x) x > 3</code></td><td>Target quantity or event indicator function h(x).</td></tr>
  <tr><td><code>f</code></td><td><code>dnorm</code></td><td>Target probability density function f(x).</td></tr>
  <tr><td><code>g</code></td><td><code>function(x) dnorm(x, mean=3)</code></td><td>Proposal probability density function g(x) centered in rare event region.</td></tr>
  <tr><td><code>rg</code></td><td><code>function(n) rnorm(n, mean=3)</code></td><td>Random sampler function for proposal density g(x).</td></tr>
  <tr><td><code>n</code></td><td><code>5000</code></td><td>Sample size.</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?importance_sampling</code> or <code>help(importance_sampling, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Step-by-Step Importance Sampling for Rare Event (X > 3) from Scratch
n <- 5000
# Target density f(x) = N(0,1)
# Proposal density g(x) = N(3,1) centered at rare event tail
target_f <- function(x) dnorm(x, mean = 0, sd = 1)
proposal_g <- function(x) dnorm(x, mean = 3, sd = 1)

# Draw candidates from proposal g(x)
x_proposal <- rnorm(n, mean = 3, sd = 1)

# Compute importance weights w(x) = f(x) / g(x)
w <- target_f(x_proposal) / proposal_g(x_proposal)
evals <- (x_proposal > 3) * w

is_estimate <- mean(evals)
is_se <- sd(evals) / sqrt(n)

cat("True P(X > 3):", 1 - pnorm(3), "
")
cat("Importance Sampling Estimate:", is_estimate, "
")
cat("Standard Error:", is_se, "
")
```

```{r practice-ex-solution}
# Step-by-Step Importance Sampling for Rare Event (X > 3) from Scratch
n <- 5000
# Target density f(x) = N(0,1)
# Proposal density g(x) = N(3,1) centered at rare event tail
target_f <- function(x) dnorm(x, mean = 0, sd = 1)
proposal_g <- function(x) dnorm(x, mean = 3, sd = 1)

# Draw candidates from proposal g(x)
x_proposal <- rnorm(n, mean = 3, sd = 1)

# Compute importance weights w(x) = f(x) / g(x)
w <- target_f(x_proposal) / proposal_g(x_proposal)
evals <- (x_proposal > 3) * w

is_estimate <- mean(evals)
is_se <- sd(evals) / sqrt(n)

cat("True P(X > 3):", 1 - pnorm(3), "
")
cat("Importance Sampling Estimate:", is_estimate, "
")
cat("Standard Error:", is_se, "
")
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think importance weights w(x) = f(x)/g(x) should be as large as possible. (Wrong! Unbounded weights cause explosive variance).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("What is the theoretical optimal proposal density g*(x) that reduces importance sampling variance to ZERO?",
    answer("g*(x) proportional to |h(x)| * f(x)", correct = ("g*(x) proportional to |h(x)| * f(x)" == "g*(x) proportional to |h(x)| * f(x)")),
    answer("g*(x) = Uniform(0,1)", correct = ("g*(x) = Uniform(0,1)" == "g*(x) proportional to |h(x)| * f(x)")),
    answer("g*(x) = Normal(0,1)", correct = ("g*(x) = Normal(0,1)" == "g*(x) proportional to |h(x)| * f(x)")),
    answer("g*(x) = f(x)^2", correct = ("g*(x) = f(x)^2" == "g*(x) proportional to |h(x)| * f(x)")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
