---
title: "Empirical CDF & Influence Functions"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🟡 Track 2: Intermediate</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 15 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Empirical Distribution Functions & Order Statistics
</div>

### 💡 Story-Driven Motivation
> When you have no theoretical model for a dataset, the Empirical CDF Fn(x) acts as the non-parametric maximum likelihood estimator of F.

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
library(srm)
# Run exported srm function:
plot(ecdf(financial_losses), main = 'Empirical CDF of Financial Losses')
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? The Glivenko-Cantelli theorem proves that Fn(x) converges uniformly to the true CDF F(x) with probability 1!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Try modifying the parameters below:
plot(ecdf(financial_losses), main = 'Empirical CDF of Financial Losses')
```

```{r practice-ex-solution}
plot(ecdf(financial_losses), main = 'Empirical CDF of Financial Losses')
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think Fn(x) is smooth. (Wrong! Fn(x) is a step function jumping by 1/n at each observed data point).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("By how much does the Empirical Cumulative Distribution Function Fn(x) step upward at each distinct data point in a sample of size n?",
    answer("1 / n", correct = ("1 / n" == "1 / n")),
    answer("1 / sqrt(n)", correct = ("1 / sqrt(n)" == "1 / n")),
    answer("n / 100", correct = ("n / 100" == "1 / n")),
    answer("0.5", correct = ("0.5" == "1 / n")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
