---
title: "Empirical CDF & Influence Functions"
description: "Construct Empirical Cumulative Distribution Functions Fn(x) and evaluate influence functions in R."
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🟡 Track 2: Intermediate</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 15 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Empirical Distribution Functions & Order Statistics
</div>

### 💡 Story-Driven Motivation
> When you have no theoretical model for a dataset, the Empirical CDF Fn(x) acts as the non-parametric maximum likelihood estimator of F.

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
# Step-by-Step Empirical CDF (EDF) Calculation from Scratch
data_obs <- c(1.2, 2.5, 3.1, 4.8, 5.2, 6.0, 7.3, 8.1)
n <- length(data_obs)

sorted_x <- sort(data_obs)
edf_y <- (1:n) / n # Step heights = 1/n

plot(sorted_x, edf_y, type = "s", xlab = "x", ylab = "Fn(x)", main = "Empirical CDF Step Function")
points(sorted_x, edf_y, pch = 16, col = "blue")
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? The Glivenko-Cantelli theorem proves that Fn(x) converges uniformly to the true CDF F(x) with probability 1!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>x</code></td><td><code>financial_losses</code></td><td>Numeric vector of observed data points.</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?ecdf</code> or <code>help(ecdf, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Step-by-Step Empirical CDF (EDF) Calculation from Scratch
data_obs <- c(1.2, 2.5, 3.1, 4.8, 5.2, 6.0, 7.3, 8.1)
n <- length(data_obs)

sorted_x <- sort(data_obs)
edf_y <- (1:n) / n # Step heights = 1/n

plot(sorted_x, edf_y, type = "s", xlab = "x", ylab = "Fn(x)", main = "Empirical CDF Step Function")
points(sorted_x, edf_y, pch = 16, col = "blue")
```

```{r practice-ex-solution}
# Step-by-Step Empirical CDF (EDF) Calculation from Scratch
data_obs <- c(1.2, 2.5, 3.1, 4.8, 5.2, 6.0, 7.3, 8.1)
n <- length(data_obs)

sorted_x <- sort(data_obs)
edf_y <- (1:n) / n # Step heights = 1/n

plot(sorted_x, edf_y, type = "s", xlab = "x", ylab = "Fn(x)", main = "Empirical CDF Step Function")
points(sorted_x, edf_y, pch = 16, col = "blue")
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think Fn(x) is smooth. (Wrong! Fn(x) is a step function jumping by 1/n at each observed data point).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("By how much does the Empirical Cumulative Distribution Function Fn(x) step upward at each distinct data point in a sample of size n?",
    answer("1 / n", correct = ("1 / n" == "1 / n")),
    answer("1 / sqrt(n)", correct = ("1 / sqrt(n)" == "1 / n")),
    answer("n / 100", correct = ("n / 100" == "1 / n")),
    answer("0.5", correct = ("0.5" == "1 / n")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
