---
title: "2D Ising Lattice Model"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🔴 Track 3: Advanced</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 15 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Lattice Neighbors & Spin Energy
</div>

### 💡 Story-Driven Motivation
> Simulating ferromagnetism and phase transitions in atomic lattices using grid neighbor flips under Boltzmann distribution exp(-beta * H).

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
# Step-by-Step 2D Ising Lattice Metropolis Model from Scratch
size <- 10
beta <- 0.44 # Inverse temperature
n_steps <- 10000

# Initialize random grid (-1 or +1)
grid <- matrix(sample(c(-1, 1), size * size, replace = TRUE), nrow = size, ncol = size)

for (step in 1:n_steps) {
  i <- sample(1:size, 1)
  j <- sample(1:size, 1)
  
  # Periodic boundary neighbors
  top <- grid[ifelse(i == 1, size, i - 1), j]
  bottom <- grid[ifelse(i == size, 1, i + 1), j]
  left <- grid[i, ifelse(j == 1, size, j - 1)]
  right <- grid[i, ifelse(j == size, 1, j + 1)]
  
  dE <- 2 * grid[i, j] * (top + bottom + left + right)
  
  if (dE <= 0 || runif(1) < exp(-beta * dE)) {
    grid[i, j] <- -grid[i, j] # Flip spin
  }
}

image(grid, col = c("black", "gold"), main = "2D Ising Spin Grid")
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? Ernst Ising solved the 1D model in his 1925 PhD thesis and incorrectly thought 2D wouldn't have phase transitions!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>size</code></td><td><code>10</code></td><td>Lattice grid dimension (size x size).</td></tr>
  <tr><td><code>beta</code></td><td><code>0.44</code></td><td>Inverse temperature parameter beta = 1/T controlling spin coupling strength.</td></tr>
  <tr><td><code>n_steps</code></td><td><code>10000</code></td><td>Total spin update attempts.</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?ising_simulation</code> or <code>help(ising_simulation, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Step-by-Step 2D Ising Lattice Metropolis Model from Scratch
size <- 10
beta <- 0.44 # Inverse temperature
n_steps <- 10000

# Initialize random grid (-1 or +1)
grid <- matrix(sample(c(-1, 1), size * size, replace = TRUE), nrow = size, ncol = size)

for (step in 1:n_steps) {
  i <- sample(1:size, 1)
  j <- sample(1:size, 1)
  
  # Periodic boundary neighbors
  top <- grid[ifelse(i == 1, size, i - 1), j]
  bottom <- grid[ifelse(i == size, 1, i + 1), j]
  left <- grid[i, ifelse(j == 1, size, j - 1)]
  right <- grid[i, ifelse(j == size, 1, j + 1)]
  
  dE <- 2 * grid[i, j] * (top + bottom + left + right)
  
  if (dE <= 0 || runif(1) < exp(-beta * dE)) {
    grid[i, j] <- -grid[i, j] # Flip spin
  }
}

image(grid, col = c("black", "gold"), main = "2D Ising Spin Grid")
```

```{r practice-ex-solution}
# Step-by-Step 2D Ising Lattice Metropolis Model from Scratch
size <- 10
beta <- 0.44 # Inverse temperature
n_steps <- 10000

# Initialize random grid (-1 or +1)
grid <- matrix(sample(c(-1, 1), size * size, replace = TRUE), nrow = size, ncol = size)

for (step in 1:n_steps) {
  i <- sample(1:size, 1)
  j <- sample(1:size, 1)
  
  # Periodic boundary neighbors
  top <- grid[ifelse(i == 1, size, i - 1), j]
  bottom <- grid[ifelse(i == size, 1, i + 1), j]
  left <- grid[i, ifelse(j == 1, size, j - 1)]
  right <- grid[i, ifelse(j == size, 1, j + 1)]
  
  dE <- 2 * grid[i, j] * (top + bottom + left + right)
  
  if (dE <= 0 || runif(1) < exp(-beta * dE)) {
    grid[i, j] <- -grid[i, j] # Flip spin
  }
}

image(grid, col = c("black", "gold"), main = "2D Ising Spin Grid")
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually forget periodic boundary conditions. (Top edge wraps to bottom, left edge wraps to right).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("What happens to the 2D Ising grid spins as inverse temperature beta increases to high values (cold temperature)?",
    answer("Spins cluster into uniform magnetic domains (+1 or -1)", correct = ("Spins cluster into uniform magnetic domains (+1 or -1)" == "Spins cluster into uniform magnetic domains (+1 or -1)")),
    answer("Spins become completely random noise", correct = ("Spins become completely random noise" == "Spins cluster into uniform magnetic domains (+1 or -1)")),
    answer("Spins all turn to zero", correct = ("Spins all turn to zero" == "Spins cluster into uniform magnetic domains (+1 or -1)")),
    answer("Grid dimensions shrink", correct = ("Grid dimensions shrink" == "Spins cluster into uniform magnetic domains (+1 or -1)")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
