---
title: "Score Function Gradient Estimator"
description: "Estimate expectation gradients using the score function log-derivative trick from scratch."
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🔴 Track 3: Advanced</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 18 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Log-Derivative Trick & Policy Gradients
</div>

### 💡 Story-Driven Motivation
> How do reinforcement learning agents (like AlphaGo) optimize expected rewards when objective functions lack explicit derivatives? The score function log-derivative trick!

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
# Step-by-Step Score Function Gradient Estimator from Scratch
n <- 5000
theta <- 1.0 # Parameter theta
h_payoff <- function(x) x # Objective reward H(X)

# 1. Draw samples X ~ N(theta, 1)
x <- rnorm(n, mean = theta, sd = 1)

# 2. Compute score function: S(x) = d/d_theta ln f(x; theta) = (x - theta)
score <- (x - theta)

# 3. Policy gradient estimate: E[ H(X) * S(X) ]
grad_est <- mean(h_payoff(x) * score)

cat("Estimated Gradient d/d_theta E[H(X)]:", grad_est, "
")
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? The Score Function method is the foundation of modern REINFORCE policy gradient algorithms in AI!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>h</code></td><td><code>function(x) x</code></td><td>Payoff or reward function H(X).</td></tr>
  <tr><td><code>theta</code></td><td><code>1</code></td><td>Distribution parameter theta.</td></tr>
  <tr><td><code>n</code></td><td><code>5000</code></td><td>Monte Carlo sample size.</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?score_function_gradient</code> or <code>help(score_function_gradient, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Step-by-Step Score Function Gradient Estimator from Scratch
n <- 5000
theta <- 1.0 # Parameter theta
h_payoff <- function(x) x # Objective reward H(X)

# 1. Draw samples X ~ N(theta, 1)
x <- rnorm(n, mean = theta, sd = 1)

# 2. Compute score function: S(x) = d/d_theta ln f(x; theta) = (x - theta)
score <- (x - theta)

# 3. Policy gradient estimate: E[ H(X) * S(X) ]
grad_est <- mean(h_payoff(x) * score)

cat("Estimated Gradient d/d_theta E[H(X)]:", grad_est, "
")
```

```{r practice-ex-solution}
# Step-by-Step Score Function Gradient Estimator from Scratch
n <- 5000
theta <- 1.0 # Parameter theta
h_payoff <- function(x) x # Objective reward H(X)

# 1. Draw samples X ~ N(theta, 1)
x <- rnorm(n, mean = theta, sd = 1)

# 2. Compute score function: S(x) = d/d_theta ln f(x; theta) = (x - theta)
score <- (x - theta)

# 3. Policy gradient estimate: E[ H(X) * S(X) ]
grad_est <- mean(h_payoff(x) * score)

cat("Estimated Gradient d/d_theta E[H(X)]:", grad_est, "
")
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually confuse score function with pathwise derivatives. (Score function works even when reward H(X) is discontinuous).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("What mathematical identity allows moving the gradient d/d_theta inside an expectation integral E_theta[H(X)]?",
    answer("d/d_theta f(x;theta) = f(x;theta) * d/d_theta log f(x;theta)", correct = ("d/d_theta f(x;theta) = f(x;theta) * d/d_theta log f(x;theta)" == "d/d_theta f(x;theta) = f(x;theta) * d/d_theta log f(x;theta)")),
    answer("Integration by parts", correct = ("Integration by parts" == "d/d_theta f(x;theta) = f(x;theta) * d/d_theta log f(x;theta)")),
    answer("Taylor expansion", correct = ("Taylor expansion" == "d/d_theta f(x;theta) = f(x;theta) * d/d_theta log f(x;theta)")),
    answer("L'Hopital's rule", correct = ("L'Hopital's rule" == "d/d_theta f(x;theta) = f(x;theta) * d/d_theta log f(x;theta)")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
