---
title: "Raw Monte Carlo Integration"
description: "Calculate Monte Carlo integrals, standard error scaling O(N^-1/2), and confidence intervals from scratch."
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🟢 Track 1: Essentials</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 12 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Expectation & Variance Bounds
</div>

### 💡 Story-Driven Motivation
> While recovering from an illness in 1946, mathematician Stanislaw Ulam played Solitaire. Instead of calculating winning odds by hand, he wondered: 'Why not play 100 games and count the wins?' Thus Monte Carlo was born.

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
# Step-by-Step Raw Monte Carlo Integration of g(x) = x^2 from Scratch
n <- 10000
a <- 0
b <- 1

# 1. Draw uniform samples U ~ Unif(a, b)
u <- runif(n, min = a, max = b)

# 2. Evaluate integrand g(u)
evals <- (b - a) * (u^2)

# 3. Compute Monte Carlo Estimate & Standard Error
mc_est <- mean(evals)
mc_se <- sd(evals) / sqrt(n)
ci_95 <- c(mc_est - 1.96 * mc_se, mc_est + 1.96 * mc_se)

cat("Monte Carlo Estimate:", mc_est, "
")
cat("Standard Error:", mc_se, "
")
cat("95% CI:", ci_95, "
")
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? The code name 'Monte Carlo' was suggested by Nicholas Metropolis after Ulam's uncle who loved the Monaco casino!
</div>


### 🔍 Function Arguments Explained

<table style="width:100%; border-collapse: collapse; margin-top:10px; margin-bottom:15px; border:1px solid #ddd;">
  <thead style="background-color:#004085; color:white;">
    <tr><th style="padding:8px; text-align:left;">Argument</th><th style="padding:8px; text-align:left;">Default</th><th style="padding:8px; text-align:left;">Description & Meaning for Students</th></tr>
  </thead>
  <tbody>
  <tr><td><code>g</code></td><td><code>function(x) x^2</code></td><td>Target function g(x) to integrate over interval [a, b].</td></tr>
  <tr><td><code>n</code></td><td><code>10000</code></td><td>Total number of uniform Monte Carlo sample points.</td></tr>
  <tr><td><code>a</code></td><td><code>0</code></td><td>Lower integration limit.</td></tr>
  <tr><td><code>b</code></td><td><code>1</code></td><td>Upper integration limit.</td></tr>
  </tbody>
</table>

<div style="background-color: #eef9f1; border-left: 5px solid #28a745; padding: 10px; margin-bottom: 15px;">
<strong>📚 Need full R Documentation while working?</strong><br/>
Run <code>?mc_integrate</code> or <code>help(mc_integrate, package = "srm")</code> in your RStudio Console to open the official package help file!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Step-by-Step Raw Monte Carlo Integration of g(x) = x^2 from Scratch
n <- 10000
a <- 0
b <- 1

# 1. Draw uniform samples U ~ Unif(a, b)
u <- runif(n, min = a, max = b)

# 2. Evaluate integrand g(u)
evals <- (b - a) * (u^2)

# 3. Compute Monte Carlo Estimate & Standard Error
mc_est <- mean(evals)
mc_se <- sd(evals) / sqrt(n)
ci_95 <- c(mc_est - 1.96 * mc_se, mc_est + 1.96 * mc_se)

cat("Monte Carlo Estimate:", mc_est, "
")
cat("Standard Error:", mc_se, "
")
cat("95% CI:", ci_95, "
")
```

```{r practice-ex-solution}
# Step-by-Step Raw Monte Carlo Integration of g(x) = x^2 from Scratch
n <- 10000
a <- 0
b <- 1

# 1. Draw uniform samples U ~ Unif(a, b)
u <- runif(n, min = a, max = b)

# 2. Evaluate integrand g(u)
evals <- (b - a) * (u^2)

# 3. Compute Monte Carlo Estimate & Standard Error
mc_est <- mean(evals)
mc_se <- sd(evals) / sqrt(n)
ci_95 <- c(mc_est - 1.96 * mc_se, mc_est + 1.96 * mc_se)

cat("Monte Carlo Estimate:", mc_est, "
")
cat("Standard Error:", mc_se, "
")
cat("95% CI:", ci_95, "
")
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually think Monte Carlo error scales as 1/N. (Wrong! By the Central Limit Theorem, error scales as 1/sqrt(N)).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("To reduce Monte Carlo standard error by a factor of 10, by how much must you increase the sample size N?",
    answer("10 times", correct = ("10 times" == "100 times")),
    answer("100 times", correct = ("100 times" == "100 times")),
    answer("1,000 times", correct = ("1,000 times" == "100 times")),
    answer("sqrt(10) times", correct = ("sqrt(10) times" == "100 times")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
