---
title: "Adaptive Metropolis (Haario et al.)"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggplot2)
library(srm)
knitr::opts_chunk$set(echo = TRUE)
gradethis::gradethis_setup()
```

## Overview & Metadata

<div style="background-color: #f0f7ff; border-left: 5px solid #0066cc; padding: 12px; margin-bottom: 15px;">
<strong>🔴 Track 3: Advanced</strong><br/>
⏱️ <strong>Estimated Time:</strong> ⏱️ 20 minutes<br/>
📋 <strong>Prerequisites:</strong> ✓ Covariance Adaptation & Ergodicity Preservation
</div>

### 💡 Story-Driven Motivation
> Manually tuning proposal covariance matrices in 50 dimensions is painful. Haario's Adaptive Metropolis dynamically updates proposal covariance Sigma_n from past MCMC history while preserving ergodic stationarity!

---

## 🎮 Interactive Playground & Visual Exploration

Experiment with the algorithm below! Change parameters and observe how the output responds:

```{r demo-code, exercise=TRUE}
library(srm)
# Run exported srm function:
adaptive_metropolis(target_log_density = function(x) -0.5 * sum(x^2), init = c(0,0), n_iter = 3000)
```

<div style="background-color: #fff8e6; border-left: 5px solid #ffb300; padding: 12px; margin-top: 15px;">
<strong>💡 Did You Know?</strong><br/>
Did You Know? Diminishing adaptation ensures the chain remains ergodic even though updates depend on past history!
</div>

---

## 🧠 Predict-First Checkpoint & Exercise

Before running the code below, predict what will happen:

```{r practice-ex, exercise=TRUE, exercise.eval=FALSE}
# Try modifying the parameters below:
adaptive_metropolis(target_log_density = function(x) -0.5 * sum(x^2), init = c(0,0), n_iter = 3000)
```

```{r practice-ex-solution}
adaptive_metropolis(target_log_density = function(x) -0.5 * sum(x^2), init = c(0,0), n_iter = 3000)
```

```{r practice-ex-check}
grade_this_code()
```

---

## ⚠️ Common Student Mistakes

<div style="background-color: #ffebe6; border-left: 5px solid #ff3300; padding: 12px; margin-bottom: 15px;">
<strong>Common Pitfall:</strong><br/>
Students usually adapt without diminishing step sizes. (Uncontrolled adaptation violates Markov property and invalidates stationarity).
</div>

---

## 📖 Mathematical Derivation (Progressive Disclosure)

<details>
<summary><strong>Click to expand formal mathematical proof & derivations</strong></summary>

### Mathematical Foundations
For full theoretical proofs, refer to Chapter 1-5 of the course notes. The primary mathematical relations are evaluated numerically by `srm` package functions.
</details>

---

## 📝 Concept Check Quiz

```{r quiz-check}
quiz(
  question("What essential condition must adaptive MCMC satisfy so that the chain still converges to the correct target distribution?",
    answer("Diminishing adaptation (adaptation steps shrink as n -> infinity)", correct = ("Diminishing adaptation (adaptation steps shrink as n -> infinity)" == "Diminishing adaptation (adaptation steps shrink as n -> infinity)")),
    answer("Constant proposal SD", correct = ("Constant proposal SD" == "Diminishing adaptation (adaptation steps shrink as n -> infinity)")),
    answer("Zero acceptance rate", correct = ("Zero acceptance rate" == "Diminishing adaptation (adaptation steps shrink as n -> infinity)")),
    answer("Linear scaling", correct = ("Linear scaling" == "Diminishing adaptation (adaptation steps shrink as n -> infinity)")),
    allow_retry = TRUE
  )
)
```

---

## 🏁 Summary & Next Steps

<div style="background-color: #e6ffe6; border-left: 5px solid #009933; padding: 12px;">
<strong>3 Key Takeaways:</strong>
1. Always understand the intuitive story before memorizing formulas.
2. Tweak parameters and monitor visual diagnostics.
3. Verify your results with R tests and package checks.
</div>
